---
title: "GII Economy Monitoring Dashboard"
output:
  officedown::rdocx_document:
    reference_docx: reference.docx
    page_margins:
      bottom: 0.5
      top: 0.5
      right: 0.5
      left: 0.5
      gutter: 0
params:
  iso3: "BTN"
  giiyr: 2025
  reported_counts_path: "reported_counts.rds"
  indicator_list_path: "full_indicator_list.rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(officer)
library(officedown)
library(flextable)

set_flextable_defaults(font.size = 8, padding = 3, theme_fun = theme_vanilla)

# ---- assumption constants (tune to match official GII methodology) --------
IN_MIN_REQUIRED   <- 35   # minimum Input indicators required (per report text)
OUT_MIN_REQUIRED  <- 15   # minimum Output indicators required (per report text)
SUBPILLAR_PASS_FRAC <- 0.5  # a sub-pillar "counts" if >= this share of its indicators are reported

# ---- palette matching the dashboard legend --------------------------------
col_new       <- "#C6E0B4"  # green
col_outdated  <- "#BDD7EE"  # blue
col_very_out  <- "#F4B183"  # orange/red (5-9y & 10y)
col_missing   <- "#F8CBAD"  # pink/red
col_excluded  <- "#D9D9D9"  # grey
col_header_in <- "#548235"  # dark green (Input pillars)
col_header_out<- "#2E75B6"  # dark blue (Output pillars)

# ---- load data --------------------------------------------------------
reported_counts <- readRDS(params$reported_counts_path)
indicator_list  <- readRDS(params$indicator_list_path)

iso3  <- params$iso3
giiyr <- params$giiyr

country_name <- iso3
if (requireNamespace("countrycode", quietly = TRUE)) {
  nm <- countrycode::countrycode(iso3, "iso3c", "country.name")
  if (!is.na(nm)) country_name <- nm
}

rc <- reported_counts %>% filter(ISO3 == iso3)
if (nrow(rc) == 0) stop("ISO3 '", iso3, "' not found in reported_counts")
rc <- rc %>% slice(1)

ind <- indicator_list %>%
  filter(ISO3 == iso3, GIIYR == giiyr) %>%
  arrange(SUBINDEX_NAME, PILLAR_NAME, SUBPILLAR, INDICATOR)
```

# GII Economy Monitoring Dashboard

## `r country_name`

This report provides an overview of the indicator availability and data timeliness for a
selected economy. For an economy to feature in the GII `r giiyr`, the minimum data
coverage requirement is at least `r IN_MIN_REQUIRED` indicators in the Innovation Input
Sub-Index and `r OUT_MIN_REQUIRED` indicators in the Innovation Output Sub-Index, with
scores for at least two sub-pillars per pillar.

```{r summary-table}
max_missing_in  <- rc$IN_INDICATORS  - IN_MIN_REQUIRED
max_missing_out <- rc$OUT_INDICATORS - OUT_MIN_REQUIRED

summary_df <- tibble(
  Scope       = c("All", "Inputs", "Outputs"),
  Total       = c(rc$N_REPORTED,  rc$IN_REPORTED,  rc$OUT_REPORTED),
  New         = c(rc$N_CURRENT,   rc$IN_CURRENT,   rc$OUT_CURRENT),
  Out_Total   = c(rc$N_OUTDATED,  rc$IN_OUTDATED,  rc$OUT_OUTDATED),
  Out_1_4     = c(rc$N_OUTDATED_1_4, rc$IN_OUTDATED_1_4, rc$OUT_OUTDATED_1_4),
  Out_5_9     = c(rc$N_OUTDATED_5_9, rc$IN_OUTDATED_5_9, rc$OUT_OUTDATED_5_9),
  Out_10      = c(rc$N_OUTDATED_10,  rc$IN_OUTDATED_10,  rc$OUT_OUTDATED_10),
  Missing     = c(rc$N_MISSING,   rc$IN_MISSING,   rc$OUT_MISSING),
  MaxMissing  = c(NA_real_, max_missing_in, max_missing_out)
)

ft <- flextable(summary_df,
  col_keys = c("Scope", "Total", "New", "Out_Total", "Out_1_4",
               "Out_5_9", "Out_10", "Missing", "MaxMissing")) %>%
  set_header_labels(
    Scope = "", Total = "Total", New = "New",
    Out_Total = "Total", Out_1_4 = "1-4y", Out_5_9 = "5-9y", Out_10 = "10y",
    Missing = "Missing", MaxMissing = "Max missing"
  ) %>%
  add_header_row(values = c("", "Reported", "", "Outdated", "", "", "", "", ""),
                  colwidths = rep(1, 9), top = TRUE) %>%
  merge_at(i = 1, j = 2:3, part = "header") %>%
  merge_at(i = 1, j = 4:7, part = "header") %>%
  align(align = "center", part = "all") %>%
  bg(j = "New", bg = col_new, part = "body") %>%
  bg(j = c("Out_Total", "Out_1_4", "Out_5_9"), bg = col_outdated, part = "body") %>%
  bg(j = "Out_10", bg = col_very_out, part = "body") %>%
  bg(j = "Missing", bg = col_missing, part = "body") %>%
  bg(j = "MaxMissing", bg = "white", part = "body") %>%
  color(j = "MaxMissing", color = "red", part = "body") %>%
  bold(j = "Scope", part = "body") %>%
  autofit()

ft
```

```{r shares-table}
pct <- function(x, n) round(100 * x / n)

shares_df <- tibble(
  Reported = pct(rc$N_REPORTED, rc$N_INDICATORS),
  New      = pct(rc$N_CURRENT,  rc$N_INDICATORS),
  Outdated = pct(rc$N_OUTDATED, rc$N_INDICATORS),
  Missing  = pct(rc$N_MISSING,  rc$N_INDICATORS)
) %>%
  mutate(across(everything(), ~ paste0(.x, "%")))

flextable(shares_df) %>%
  align(align = "center", part = "all") %>%
  fontsize(size = 14, part = "body") %>%
  autofit()
```

```{r pillar-constraint}
# A sub-pillar "counts" toward the pillar total if at least SUBPILLAR_PASS_FRAC
# of its indicators are reported (assumption - tune constant above).
subpillar_status <- ind %>%
  group_by(SUBINDEX_NAME, PILLAR_NAME, SUBPILLAR) %>%
  summarise(
    n_ind     = n(),
    n_reported = sum(!is.na(DATAYR)),
    pass = (n_reported / n_ind) >= SUBPILLAR_PASS_FRAC,
    .groups = "drop"
  )

pillar_order <- c("Institutions", "Human capital and research", "Infrastructure",
                   "Market sophistication", "Business sophistication",
                   "Knowledge and technology outputs", "Creative outputs")

pillar_df <- subpillar_status %>%
  group_by(PILLAR_NAME) %>%
  summarise(n_subpillars_passing = sum(pass), .groups = "drop") %>%
  right_join(tibble(PILLAR_NAME = pillar_order), by = "PILLAR_NAME") %>%
  mutate(n_subpillars_passing = replace_na(n_subpillars_passing, 0)) %>%
  arrange(match(PILLAR_NAME, pillar_order))

is_input_pillar <- pillar_df$PILLAR_NAME %in%
  c("Institutions", "Human capital and research", "Infrastructure",
    "Market sophistication", "Business sophistication")

row_names  <- as_tibble(setNames(as.list(pillar_df$PILLAR_NAME), pillar_order))
row_counts <- as_tibble(setNames(as.list(as.character(pillar_df$n_subpillars_passing)), pillar_order))
pc_wide <- bind_rows(row_names, row_counts)

ft_pc <- flextable(pc_wide) %>%
  delete_part(part = "header") %>%
  align(align = "center", part = "all") %>%
  fontsize(size = 7, part = "body") %>%
  bold(i = 2) %>%
  autofit()

for (k in seq_along(pillar_order)) {
  hdr_col <- if (is_input_pillar[k]) col_header_in else col_header_out
  ft_pc <- bg(ft_pc, i = 1, j = k, bg = hdr_col, part = "body")
  ft_pc <- color(ft_pc, i = 1, j = k, color = "white", part = "body")
  ft_pc <- bg(ft_pc, i = 2, j = k, bg = col_new, part = "body")
}

ft_pc
```

\newpage

## Full indicator reporting status for GII `r giiyr`

```{r indicator-status}
ind_status <- ind %>%
  mutate(
    excluded = str_detect(NAME, "\\*$"),
    age = giiyr - DATAYR,
    status = case_when(
      excluded            ~ "EXCLUDED",
      is.na(DATAYR)       ~ "MISSING",
      age <= 0             ~ "NEW",
      age >= 1 & age <= 4  ~ "OUTDATED_1_4",
      age >= 5 & age <= 9  ~ "OUTDATED_5_9",
      age >= 10            ~ "OUTDATED_10",
      TRUE                 ~ "NEW"
    )
  ) %>%
  select(SUBINDEX_NAME, PILLAR_NAME, SUBPILLAR, INDICATOR, NAME, DATAYR, MODEYR, status)

status_colors <- c(
  NEW = col_new, OUTDATED_1_4 = col_outdated, OUTDATED_5_9 = col_outdated,
  OUTDATED_10 = col_very_out, MISSING = col_missing, EXCLUDED = col_excluded
)

ft_ind <- flextable(ind_status,
  col_keys = c("SUBINDEX_NAME", "PILLAR_NAME", "SUBPILLAR", "INDICATOR",
               "NAME", "DATAYR", "MODEYR")) %>%
  set_header_labels(
    SUBINDEX_NAME = "SubIdx", PILLAR_NAME = "Pillar", SUBPILLAR = "SubPillar",
    INDICATOR = "Level", NAME = "Indicator Name",
    DATAYR = "Data Year", MODEYR = "Mode Year"
  ) %>%
  fontsize(size = 7, part = "all") %>%
  align(j = c("DATAYR", "MODEYR"), align = "center", part = "all") %>%
  autofit()

for (s in names(status_colors)) {
  rows <- which(ind_status$status == s)
  if (length(rows) > 0) {
    ft_ind <- bg(ft_ind, i = rows, bg = status_colors[[s]], part = "body")
  }
}

ft_ind
```

*Indicator is excluded from the data minimum criteria (DMC) rules. Legend: NEW (green),
OUTDATED (blue), VERY OUTDATED (orange), MISSING (pink), EXCLUDED (grey).
